---
title: "atomicvectors_2024_02_01"
author: "Emily Dombrowski"
date: "2024-02-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Comp bio 2024-02-1
Markdown for evaluating atomic vectors and useful functions
```{r}
x <- 5 # preferred
y = 4 # legal but not used except in function defaults
y = y + 1.1
print(y)
```

```{r}
y <- y + 1.1
print(y)
```

# Variable names
```{r}
z <- 3 # begin with lower case. letter
plantHeight <- 10 # option "camelCaseFormatting"
plant.height <- 4.2 # acoid periods
plant_height <- 3.3 # optimal "snake_case_formatting"
. <- 5.5 # reserve this for a generic temporary variable (more later)
print(.) # this works but its not showing up in my environment. Kind of weird
```

# Types of variables
* Characters
* Doubles
* Factors (characters but used for statistical factors)
* Integers
* Vectors with lists

# Combine function
```{r}
z <- c(3.2, 5, 5, 6)
print(z)
typeof(z)
is.numeric(z)
```

## (c) always "flattens" to an atomic vector
Flatten means it doesn't retain structure of concatenation. It ignores internal structure and strings together functions. 
```{r}
z <- c(c(3,4), c(5,6))
print(z)
```

# Building character strings with single or double quotes
```{r}
z <- c("perch", "bass", 'trout')
# flexibility with character strings single or double quotes
print(z)

z <- c("This is only 'one' character string", 'a second')
print(z)
typeof(z)
is.character(z)
```

# Building logicals
```{r}
# Boolean, not with quotes, all caps
z <- c(TRUE,TRUE,FALSE)
# avoid abbreviations T, F which will work
print(z)
typeof(z)
is.logical(z)
is.integer(z)
```

# Specifing elements of a matrix
```{r}
dogs <- c("chow","pug","beagle","greyhound","akita")
dogs[1] # picks first element
dogs[2] 
dogs[6] # outside of bounds, gives us [NA]

# pass the brackets a group of elements (= a vector) to subset the vector
dogs[c(3,5)]

# works with multiple repeats
dogs[c(1,1,1,4)]
dogs[] # gives us the whole df

length(dogs)
dogs[5]
dogs[length(dogs)]

# sometimes, it is faster to specify what we *don't want* in vectors.
dogs[-1]

# we can also use double negatives
dogs[c(-2,-4)]
dogs[-c(2,4)]
dogs[c(-1,2)] # gives error, doesn't want different operations (+/-) in same command
```
While coding, using numbers in this way isn't super useful--if one line of a file changes, it will break your whole code. Using objects and functions is better.

# Three properties of vectors
## Type
```{r}
z <- c(1.1, 1.2, 3, 4.4)
typeof(z)
is.numeric(z)
as.character(z) # doesn't change z permanantly/globally
print(z)
typeof(z)

z2 <- as.character(z)
print(z2)
typeof(z2)

length(z)
length(y_not) # gives an error if variable doesn't exist
```

## Names
```{r}
z <- runif(5)
# optional attribute not initially assigned
print(z)
names(z)
names(z) <- c("chow","pug","beagle","greyhound","akita")
print(z)

z2 <- c(gold=3.3, silver=10, lead=2)
print(z2)

# reset names
names(z2) <- NULL
print(z2)
names(z2) <- c("COPPER", "ZINC")
```

# Special Data Types
```{r}
# NA values
z <- c(3.2, 3.3, NA)
typeof(z)
length(z)
typeof(z[3]) # keeps track of variable as double

z1 <- NA
typeof(z1)

is.na(z) # output is boolean. Logical to find missing values
!is.na(z)
mean(z)
```

```{r}
a <- c(2, 2.0 )
typeof(a)

# if elements are of different types r coerces them to a
# common type...hierarchy
# logical -> integer -> double -> character

a <-c(2, 3, "3")
print(a)
# converts a to character

a <- c(1,2,2,l)
print(a)
# gives error and allows us to identify issues in data
# collection

a <- runif(10)
print(a)

# comparison operators yield a logical result
a > 0.50

# conduct math operations on this vector of logical
# tru/fales values

# how many elements in the vector are > 0.5?
sum(TRUE, TRUE) # value gets coerced to integer
sum(a > 0.5) 

# what proportion of vector values are > 0.05?
sum(a > 0.5)/length(a)
mean(a > 0.5)

# qualifying exam question in biology
# approximately what proportion of observations drawn from
# a standard normal distribution are larger than 2.0

mean(rnorm(1,000) > 2.0) 
```
# vectorization
```{r}
z <- c(10,20,30)
z + 1
z <- 10
z + 1

z <- c(10, 20, 30)
y <- c(1, 2, 3)

z + y

y^2

10 + 3*y + 2*y^2
# applies to each value in the list
```
Powerful operaations compactly by setting it up this way. This makes r coding pretty compact, in other languages we'd need to write for loops to address this.

# Recycling
What if the vector lengths are not equal to eachother?
```{r}
z <- c(10,20,30)
x <- c(1,2)

z+x
```
We get an output alongside an error message: "Warning: longer object length is not a multiple of shorter object length." It goes back and recycles the 1 and the first value in the shorter vector gets used again. 