---
title: "hw9.Rmd"
author: "Emily Dombrowski"
date: "2024-03-27"
output: html_document
---
# "Script" 1: Simulate Data. 
This is a back-end function file that is from an R script. This file is used to have more extensive documentation about functions and to completely generate everything before it goes into the final script. 
```{r}
# Data generation file
# FunctionsFile.R

library(ggplot2)

# global variables for df input
mean_vec = c(0.65, 0.85, 0.37, 0.71, 0.52, 0.75)
sd_vec = c(0.048, 0.076, 0.086, 0.053, 0.041, 0.076)
num = 50

##################################################
# function: gendata
# This is a function to generate simulated data using rnorm for multiple
# parameters, then to create a dataframe.
# input: number of observations, mean values, and standard deviations
# output: a data frame for analysis
#-------------------------------------------------


gendata <- function(n = num, mean = mean_vec, sd = sd_vec){
  mg_sur_25 <- rnorm(n = num, mean = mean_vec[1],
                     sd = sd_vec[1])
  
  mg_sur_18 <- rnorm(n = num, mean = mean_vec[2],
                     sd = sd_vec[2])
  
  mp_sur_25 <- rnorm(n = num, mean = mean_vec[3],
                     sd = sd_vec[3])
  
  mp_sur_18 <- rnorm(n = num, mean = mean_vec[4],
                     sd = sd_vec[4])
  
  hy_sur_25 <- rnorm(n = num, mean = mean_vec[5],
                     sd = sd_vec[5])
  
  hy_sur_18 <- rnorm(n = num, mean = mean_vec[6],
                     sd = sd_vec[6])
  
  conditions <- c(mg_sur_25, mg_sur_18, mp_sur_18,
                  mp_sur_18,hy_sur_25, hy_sur_18)
  
  temp_treatment <- c(rep(18, 150), rep(25, 150))
  
  species <- c(rep("MP", 50), rep("MG", 50), rep("HY", 50),rep("MP", 50), rep("MG", 50), rep("HY", 50))
  
  aov_df <- data.frame(index=1:length(conditions),
                       surv_rate=conditions,
                       temp=temp_treatment,
                       species=species)
  return(aov_df)
}


#######################
# function: sumstats
# This is a function to generate summary statistics from the previously 
# generated dataframe.
# input: previously generated df
# output: summary table
#-------------------------------------------------

sumstats <- function(working_df = NULL){
  working_df = working_df[2]
  sum_stats <- unlist(summary(working_df))
  return(sum_stats)
}

aov_mod <- function(working_df = NULL){
  anoModel <- aov(surv_rate~temp + species + temp:species,
                  data=working_df)
  sumstats <- summary(anoModel)
  return(sumstats)
}

# aov_mod(gendata()) # test for debugging


#######################
# function: plot_func
# This is a function to generate summary plots from the previously 
# generated dataframe.
# input: dataframe
# output: plots
#-------------------------------------------------
plot_func <- function(working_df = NULL){
  boxplot <-ggplot(data=working_df) + 
    aes(x=as.factor(temp), 
        y=surv_rate, 
        fill=species) + 
    geom_boxplot() + 
    xlab("Temperature (°C)") + ylab("Survival Rate")
  
  plot <- boxplot + labs(fill = "Species") + 
    theme(panel.background = element_blank())

return(plot)
  print(plot)
}
 
```
# Script 2: running the code in a few lines
This "front-end" file concisely carries out the relevant operations in very few lines of code by sourcing the file.
```{r}
# EGD
# Homework 9
# Front-end file 
source("FunctionsFile.R")

x <- gendata()
y <- plot_func(x)
z <- sumstats(x)

print(head(x))
print(y)
print(z)
```
