---
title: "Working with Markdown"
author: "Emily Dombrowski"
date: "2024-01-23"
output: 
  html_document: 
    theme: yeti
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is a practice document with markdown formatting and practice LaTex commands. 

For markdown, use asterisks to *indicate italic text*.

Double line return for a new chunk of text like this. 

Emily Dombrowski  
Department of Biology  
University of Vermont 

# Hash marks for headers

## More hashes for progressively

### smaller

#### and smaller

#### headers!

**Boldface font with double asterisk**

>Offset quote which embeds and looks really nice when you have extended material that you would like to see.

* first list item
  * Item 2
  
^superscripts^ but also ~subscripts~ like this

[set up a link with text](website address)

also footnotes^[at bottom of page]

First header | Second header
-------------|--------------
Content cell | Content cell
Content cell | Content cell

```{r}
# first  chunk of code in rmd file
print(pi)
pi + 3
```

```
Fencing sets boundaries and everything inside the fencing *is shown as literal text* without rendering
```

$$a + b =c$$

We can also use Latex in line with a single dollar sign $a +b=c$ like this.

## Subscripts in LaTex

$$ H_0 = Z_{a +b}$$


$$S=cA^z_1 + z_{a+b}$$

## Fractions and greek letters
$${\alpha = \frac{\beta}{\delta + \gamma_x}}$$

$$z + \sum_{i=1}^X{k}$$

## Escaping the backslash character
$$\backslash \alpha \le b \backslash$$

## Rendering in plain text
$$P(Occurrence of Species A) = Z$$ 
...looks wrong! Renders plain text as mathmatical variables. 

$$P(\mbox{Occurrence of Species A}) = Z$$
Used mbox command which alleviates this issue

Easy to embed latex in line $\pi + 3$ like so. We can also use markdown fencing in line with a single backtick `pi + 3` like this. We can even execute r code inside of markdown text `r pi + 3` (reads out as computed answer rather than unformatted code.)

```{r eval=TRUE, echo=FALSE}
# Use comments extensively in ALL of your coding!
Pred <- seq(1,10)     # make a vector of integers from 1 to 10
Res <- runif(10)      # generate 10 draws from a random uniform (0,1) distribution

# print the random numbers
print(Res)

# plot the graph
plot(x=Pred,y=Res,type="b")
```
```{r}
plot(runif(50),cex=10*runif(50),col="goldenrod")

```